### 2019-09-20 ch8_guidelines/content.tex
markup changes

Benchmarking studies are carried out by computational researchers to compare the performance of different methods, using reference datasets and a range of evaluation criteria. Benchmarks may be performed by authors of new methods to demonstrate performance improvements or other advantages; by independent groups interested in systematically comparing existing methods; or organized as community challenges. 'Neutral' benchmarking studies, i.e., those performed independently of new method development by authors without any perceived bias, and with a focus on the comparison itself, are especially valuable for the research community \cite{boulesteix_necessitydesignstudies_2018,boulesteix_pleaneutralcomparison_2013}.

@@

Our aim is to complement previous reviews by providing a summary of essential guidelines for designing, performing, and interpreting benchmarks. While all guidelines are essential for a truly excellent benchmark, some are more fundamental than others. Our target audience consists of computational researchers who are interested in performing a benchmarking study, or who have already begun one. Our review spans the full 'pipeline' of benchmarking, from defining the scope to best practices for reproducibility. This includes crucial questions regarding design and evaluation principles: for example, using rankings according to evaluation metrics to identify a set of high-performing methods, and then highlighting different strengths and tradeoffs among these.

@@

The purpose and scope of a benchmark should be clearly defined at the beginning of the study, and will fundamentally guide the design and implementation. In general, we can define three broad types of benchmarking studies: (i) those by method developers, to demonstrate the merits of their approach (e.g., \cite{weber_diffcytdifferentialdiscovery_2019,nowicka_drimseqdirichletmultinomialframework_2016,levine_datadrivenphenotypicdissection_2015,zhou_robustlydetectingdifferential_2014,law_voomprecisionweights_2014}); (ii) neutral studies performed to systematically compare methods for a certain analysis, either conducted directly by an independent group (e.g., \cite{saelens_comparisonsinglecelltrajectory_2019,duo_systematicperformanceevaluation_2018,soneson_biasrobustnessscalability_2018,saelens_comprehensiveevaluationmodule_2018,weber_comparisonclusteringmethods_2016,korthauer_practicalguidemethods_2019,freytag_comparisonclusteringtools_2018,baruzzo_simulationbasedcomprehensivebenchmarking_2017,kanitz_comparativeassessmentmethods_2015,soneson_comparisonmethodsdifferential_2013,rapaport_comprehensiveevaluationdifferential_2013,dillies_comprehensiveevaluationnormalization_2013}) or in collaboration with method authors (e.g., \cite{sage_quantitativeevaluationsoftware_2015}); or (iii) those organized in the form of a community challenge, such as those from the DREAM \cite{weirauch_evaluationmethodsmodeling_2013,costello_communityeffortassess_2014,kuffner_crowdsourcedanalysisclinical_2015,ewing_combiningtumorgenome_2015,hill_inferringcausalmolecular_2016}, FlowCAP \cite{aghaeepour_criticalassessmentautomated_2013,aghaeepour_benchmarkevaluationalgorithms_2016}, CASP \cite{moult_criticalassessmentmethods_2018,moult_criticalassessmentmethods_2016}, CAMI \cite{sczyrba_criticalassessmentmetagenome_2017}, Assemblathon \cite{earl_assemblathoncompetitiveassessment_2011,bradnam_assemblathonevaluatingnovo_2013}, MAQC/SEQC \cite{maqcconsortium_microarrayqualitycontrol_2006,shi_microarrayqualitycontrol_2010,seqc_comprehensiveassessmentrnaseq_2014}, and GA4GH \cite{krusche_bestpracticesbenchmarking_2019} consortia.

@@

By contrast, when introducing a new method, the focus of the benchmark will be on evaluating the relative merits of the new method. This may be sufficiently achieved with a less extensive benchmark, e.g., by comparing against a smaller set of state-of-the-art and baseline methods. However, the benchmark must still be carefully designed to avoid disadvantaging any methods; for example, extensively tuning parameters for the new method while using default parameters for competing methods would result in a biased representation. Some advantages of a new method may fall outside the scope of a benchmark; for example, a new method may enable more flexible analyses than previous methods (e.g., beyond two-group comparisons in differential analyses \cite{weber_diffcytdifferentialdiscovery_2019}).

@@

The selection of methods to include in the benchmark will be guided by the purpose and scope of the study. A neutral benchmark should include all available methods for a certain type of analysis. In this case, the publication describing the benchmark will also function as a review of the literature; a summary table describing the methods is a key output (e.g., Figure 2 in \cite{saelens_comparisonsinglecelltrajectory_2019} or Table 1 in \cite{weber_comparisonclusteringmethods_2016}). Alternatively, it may make sense to include only a subset of methods, by defining inclusion criteria: for example, all methods that (i) provide freely available software implementations, (ii) are available for commonly used operating systems, and (iii) can successfully be installed without errors following a reasonable amount of trouble-shooting. Such criteria should be chosen without favoring any methods, and exclusion of any widely used methods should be justified. A useful strategy can be to involve method authors within the process, since they may provide additional details on optimal usage. In addition, community involvement can lead to new collaborations and inspire future method development. However, the overall neutrality and balance of the resulting research team should be maintained. Finally, if the benchmark is organized as a community challenge, the selection of methods will be determined by the participants. In this case, it is important to communicate the initiative widely -- for example, through an established network such as DREAM challenges. However, some authors may choose not to participate; a summary table documenting non-included methods should be provided in this case.

When developing a new method, it is generally sufficient to select a representative subset of existing methods to compare against. For example, this could consist of the current best-performing methods (if known), a simple 'baseline' method, and any methods that are widely used. The selection of competing methods should ensure an accurate and unbiased assessment of the relative merits of the new approach, compared with the current state-of-the-art. In fast-moving fields, for a truly excellent benchmark, method developers should be prepared to update their benchmarks or design them to easily allow extensions as new methods emerge.

@@

Simulated data have the advantage that a known true signal (or 'ground truth') can easily be introduced; for example, whether a gene is differentially expressed. Quantitative performance metrics measuring the ability to recover the known truth can then be calculated. However, it is important to demonstrate that simulations accurately reflect relevant properties of real data, by inspecting empirical summaries of both simulated and real datasets (e.g., using automated tools \cite{soneson_unifiedqualityverification_2018}). The set of empirical summaries to use is context-specific; for example, for single-cell RNA-sequencing, dropout profiles and dispersion-mean relationships should be compared \cite{soneson_biasrobustnessscalability_2018}; for DNA methylation, correlation patterns among neighboring CpG sites should be investigated \cite{korthauer_detectionaccuratefalse_2019}; for comparing mapping algorithms, error profiles of the sequencing platforms should be considered \cite{caboche_comparisonmappingalgorithms_2014}. Simplified simulations can also be useful, to evaluate a new method under a basic scenario, or to systematically test aspects such as scalability and stability. However, overly simplistic simulations should be avoided, since these will not provide useful information on performance. A further advantage of simulated data is that it is possible to generate as much data as required; for example, to study variability and draw statistically valid conclusions.

Experimental data often do not contain a ground truth, making it difficult to calculate performance metrics. Instead, methods may be evaluated by comparing them against each other (e.g., overlap between sets of detected differential features \cite{nowicka_drimseqdirichletmultinomialframework_2016}), or against a current widely accepted method or 'gold standard' (e.g., manual gating to define cell populations in high-dimensional cytometry \cite{weber_comparisonclusteringmethods_2016, aghaeepour_criticalassessmentautomated_2013}, or fluorescence in situ hybridization to validate absolute copy number predictions \cite{zheng_benchmarkingcontextsdetails_2017}). In the context of supervised learning, the response variable to be predicted is known in the manually labeled training and test data. However, individual datasets should not be overused, and using the same dataset for both method development and evaluation should be avoided, due to the risk of overfitting and overly optimistic results \cite{grimm_evaluationtoolsused_2015,jelizarow_overoptimismbioinformaticsillustration_2010}. In some cases, it is also possible to design experimental datasets containing a ground truth. Examples include: (i) 'spiking in' synthetic RNA molecules at known relative concentrations \cite{jiang_syntheticspikeinstandards_2011} in RNA-sequencing experiments (e.g., \cite{seqc_comprehensiveassessmentrnaseq_2014,garalde_highlyparalleldirect_2018}), (ii) large-scale validation of gene expression measurements by quantitative polymerase chain reaction (e.g., \cite{seqc_comprehensiveassessmentrnaseq_2014}), (iii) using genes located on sex chromosomes as a proxy for silencing of DNA methylation status (e.g., \cite{law_voomprecisionweights_2014,fang_genomiclandscapehuman_2012}), (iv) using fluorescence-activated cell sorting to sort cells into known subpopulations prior to single-cell RNA-sequencing (e.g., \cite{soneson_biasrobustnessscalability_2018, schaum_singlecelltranscriptomics20_2018, zheng_massivelyparalleldigital_2017}), or (v) mixing different cell lines to create 'pseudo-cells' \cite{tian_benchmarkingsinglecell_2019}. However, it may be difficult to ensure that the ground truth represents an appropriate level of variability -- for example, the variability of spiked-in material, or whether method performance on cell line data is relevant to outbred populations. Alternatively, experimental datasets may be evaluated qualitatively, for example, by judging whether each method can recover previous discoveries, although this strategy relies on the validity of previous results.

A further technique is to design 'semi-simulated' datasets that combine real experimental data with an 'in silico' (i.e., computational) spike-in signal; for example, by combining cells or genes from 'null' (e.g., healthy) samples with a subset of cells or genes from samples expected to contain a true differential signal (examples include \cite{weber_diffcytdifferentialdiscovery_2019,arvaniti_sensitivedetectionrare_2017,rigaill_syntheticdatasets_2018}). This strategy can create datasets with more realistic levels of variability and correlation, together with a ground truth.

Overall, there is no perfect reference dataset, and the selection of appropriate datasets will involve tradeoffs, e.g., regarding the level of complexity. Both simulated and experimental data should not be too 'simple' (e.g., two of the datasets in the FlowCAP-II challenge \cite{aghaeepour_criticalassessmentautomated_2013} gave perfect performance for several algorithms) or too 'difficult' (e.g., for the third dataset in FlowCAP-II, no algorithms performed well); in these situations, it can be impossible to distinguish performance. In some cases, individual datasets have also been found to be unrepresentative, leading to over-optimistic or otherwise biased assessment of methods (e.g., \cite{lowes_bralibasedentatale_2017}). Overall, the key to truly excellent benchmarking is diversity of evaluations, i.e., using a range of metrics and datasets that span the range of those that might be encountered in practice, so that performance estimates can be credibly extrapolated.

@@

There are three major strategies for choosing parameters. The first (and simplest) is to use default values for all parameters. Default parameters may be adequate for many methods, although this is difficult to judge in advance. While this strategy may be viewed as too simplistic for some neutral benchmarks, it reflects typical usage. We used default parameters in several neutral benchmarks where we were interested in performance for untrained users \cite{saelens_comparisonsinglecelltrajectory_2019,couronne_randomforestlogistic_2018,schneider_mortalityriskacute_2016}. In addition, for \cite{saelens_comparisonsinglecelltrajectory_2019}, due to the large number of methods and datasets, total runtime was already around a week using 192 processor cores, necessitating judgment in the scope of parameter tuning. The second strategy is to choose parameters based on previous experience or published values. This relies on familiarity with the methods and the literature, reflecting usage by expert users. The third strategy is to use a systematic or automated parameter tuning procedure -- for example, a 'grid search' across ranges of values for multiple parameters or techniques such as cross-validation (e.g., \cite{saelens_comprehensiveevaluationmodule_2018}). The strategies may also be combined, e.g., setting non-critical parameters to default values and performing a grid search for key parameters. Regardless, neutrality should be maintained: comparing methods with the same strategy makes sense, while comparing one method with default parameters against another with extensive tuning makes for an unfair comparison.

For benchmarks performed to introduce a new method, comparing against a single set of optimal parameter values for competing methods is often sufficient; these values may be selected during initial exploratory work or by consulting documentation. However, as outlined above, bias may be introduced by tuning the parameters of the new method more extensively. The parameter selection strategy should be transparently discussed during the interpretation of the results, to avoid the risk of over-optimistic reporting due to expending more 'researcher degrees of freedom' on the new method \cite{boulesteix_tensimplerules_2015,hu_parametertuningkey_2019}.

Software versions can also influence results, especially if updates include major changes to methodology (e.g., \cite{vaquero-garcia_leafcuttervsmajiq_2018}). Final results should generally be based on the latest available versions, which may require re-running some methods if updates become available during the course of a benchmark.

@@

Evaluation of methods will rely on one or more quantitative performance metrics (Figure  \ref{fig:fig2_metrics}A). The choice of metric depends on the type of method and data. For example, for classification tasks with a ground truth, metrics include the true positive rate (TPR; sensitivity or recall), false positive rate (FPR; 1 - specificity), and false discovery rate (FDR). For clustering tasks, common metrics include the F1 score, adjusted Rand index, normalized mutual information, precision, and recall; some of these can be calculated at the cluster level as well as averaged (and optionally weighted) across clusters (e.g., these metrics were used to evaluate clustering methods in our own work \cite{duo_systematicperformanceevaluation_2018,weber_comparisonclusteringmethods_2016}  and by others \cite{freytag_comparisonclusteringtools_2018,aghaeepour_criticalassessmentautomated_2013,wiwie_comparingperformancebiomedical_2015}). Several of these metrics can also be compared visually to capture the tradeoff between sensitivity and specificity, e.g., using receiver operating characteristic (ROC) curves (TPR versus FPR), TPR versus FDR curves, or precision-recall (PR) curves (Figure \ref{fig:fig2_metrics}B). For imbalanced datasets, PR curves have been shown to be more informative than ROC curves \cite{saito_precisionrecallplotmore_2015,powers_visualizationtradeoffevaluation_2015}. These visual metrics can also be summarized as a single number, such as area under the ROC or PR curve; examples from our work include \cite{weber_diffcytdifferentialdiscovery_2019,soneson_biasrobustnessscalability_2018}. In addition to the tradeoff between sensitivity and specificity, a method's 'operating point' is important; in particular, whether the threshold used (e.g., 5\% FDR) is calibrated to achieve the specified error rate. We often overlay this onto TPR-FDR curves by filled or open circles (e.g., Figure \ref{fig:fig2_metrics}B, generated using the iCOBRA package \cite{soneson_icobraopenreproducible_2016}); examples from our work include \cite{weber_diffcytdifferentialdiscovery_2019,nowicka_drimseqdirichletmultinomialframework_2016,zhou_robustlydetectingdifferential_2014,soneson_differentialanalysesrnaseq_2016}.

@@

For methods with continuous-valued output (e.g., effect sizes or abundance estimates), metrics include the root mean square error, distance measures, Pearson correlation, sum of absolute log-ratios, log-modulus, and cross-entropy. As above, the choice of metric depends on the type of method and data (e.g., \cite{costello_communityeffortassess_2014,lindgreen_evaluationaccuracyspeed_2016}  used correlation, while \cite{moult_criticalassessmentmethods_2016} used root mean square deviation). Further classes of methods include those generating graphs, phylogenetic trees, overlapping clusters, or distributions; these require more complex metrics. In some cases, custom metrics may need to be developed (e.g., we defined new metrics for topologies of developmental trajectories in \cite{saelens_comparisonsinglecelltrajectory_2019}). When designing custom metrics, it is important to assess their reliability across a range of prediction values (e.g., \cite{gurevich_quastqualityassessment_2013,narzisi_comparingnovogenome_2011}). For some metrics, it may also be useful to assess uncertainty, e.g., via confidence intervals. In the context of supervised learning, classification or prediction accuracy can be evaluated by cross-validation, bootstrapping, or on a separate test dataset (e.g., \cite{gardner_identifyingaccuratemetagenome_2019,aghaeepour_benchmarkevaluationalgorithms_2016}). In this case, procedures to split data into training and test sets should be appropriate for the data structure and the prediction task at hand (e.g., leaving out whole samples or chromosomes \cite{schreiber_pitfallmachinelearning_2019}).

Additional metrics that do not rely on a ground truth include measures of stability, stochasticity, and robustness. These measures may be quantified by running methods multiple times using different inputs or subsampled data (e.g., we observed substantial variability in performance for some methods in \cite{soneson_biasrobustnessscalability_2018,weber_comparisonclusteringmethods_2016}). 'Missing values' may occur if a method does not return any values for a certain metric, e.g., due to a failure to converge or other computational issues such as excessive runtime or memory requirements (e.g., \cite{saelens_comparisonsinglecelltrajectory_2019, soneson_biasrobustnessscalability_2018,weber_comparisonclusteringmethods_2016}). Fallback solutions such as imputation may be considered in this case \cite{bischl_benchmarkinglocalclassification_2013}, although these should be transparently reported. For non-deterministic methods (e.g., with random starts or stochastic optimization), variability in performance when using different random seeds or subsampled data should be characterized. Null comparisons can be constructed by randomizing group labels such that datasets do not contain any true signal, which can provide information on error rates (e.g., \cite{weber_diffcytdifferentialdiscovery_2019,zhou_robustlydetectingdifferential_2014,law_voomprecisionweights_2014}). However, these must be designed carefully to avoid confounding by batch or population structure, and to avoid strong within-group batch effects that are not accounted for.

For most benchmarks, multiple metrics will be relevant. Focusing on a single metric can give an incomplete view: methods may not be directly comparable if they are designed for different tasks, and different users may be interested in different aspects of performance. Therefore, a crucial design decision is whether to focus on an overall ranking, e.g., by combining or weighting multiple metrics. In general, it is unlikely that a single method will perform best across all metrics, and performance differences between top-ranked methods for individual metrics can be small. Therefore, a good strategy is to use rankings from multiple metrics to identify a set of consistently high-performing methods, and then highlight the different strengths of these methods. For example, in \cite{weber_comparisonclusteringmethods_2016}, we identified methods that gave good clustering performance, and then highlighted differences in runtimes among these. In several studies, we have presented results in the form of a graphical summary of performance according to multiple criteria (examples include Figure 3 in \cite{saelens_comparisonsinglecelltrajectory_2019} and Figure 5 in \cite{soneson_biasrobustnessscalability_2018} from our work; and Figure 2 in \cite{sage_quantitativeevaluationsoftware_2015} and Figure 6 in \cite{korthauer_practicalguidemethods_2019} from other authors). Identifying methods that consistently underperform can also be useful, to allow readers to avoid these.

>>>>

Benchmarking studies are carried out by computational researchers to compare the performance of different methods, using reference datasets and a range of evaluation criteria. Benchmarks may be performed by authors of new methods to demonstrate performance improvements or other advantages; by independent groups interested in systematically comparing existing methods; or organized as community challenges. Neutral benchmarking studies, i.e., those performed independently of new method development by authors without any perceived bias, and with a focus on the comparison itself, are especially valuable for the research community \cite{boulesteix_necessitydesignstudies_2018,boulesteix_pleaneutralcomparison_2013}.

@@

Our aim is to complement previous reviews by providing a summary of essential guidelines for designing, performing, and interpreting benchmarks. While all guidelines are essential for a truly excellent benchmark, some are more fundamental than others. Our target audience consists of computational researchers who are interested in performing a benchmarking study, or who have already begun one. Our review spans the full pipeline of benchmarking, from defining the scope to best practices for reproducibility. This includes crucial questions regarding design and evaluation principles: for example, using rankings according to evaluation metrics to identify a set of high-performing methods, and then highlighting different strengths and tradeoffs among these.

@@

The purpose and scope of a benchmark should be clearly defined at the beginning of the study, and will fundamentally guide the design and implementation. In general, we can define three broad types of benchmarking studies: (i) those by method developers, to demonstrate the merits of their approach (e.g. \cite{weber_diffcytdifferentialdiscovery_2019,nowicka_drimseqdirichletmultinomialframework_2016,levine_datadrivenphenotypicdissection_2015,zhou_robustlydetectingdifferential_2014,law_voomprecisionweights_2014}); (ii) neutral studies performed to systematically compare methods for a certain analysis, either conducted directly by an independent group (e.g. \cite{saelens_comparisonsinglecelltrajectory_2019,duo_systematicperformanceevaluation_2018,soneson_biasrobustnessscalability_2018,saelens_comprehensiveevaluationmodule_2018,weber_comparisonclusteringmethods_2016,korthauer_practicalguidemethods_2019,freytag_comparisonclusteringtools_2018,baruzzo_simulationbasedcomprehensivebenchmarking_2017,kanitz_comparativeassessmentmethods_2015,soneson_comparisonmethodsdifferential_2013,rapaport_comprehensiveevaluationdifferential_2013,dillies_comprehensiveevaluationnormalization_2013}) or in collaboration with method authors (e.g. \cite{sage_quantitativeevaluationsoftware_2015}); or (iii) those organized in the form of a community challenge, such as those from the DREAM \cite{weirauch_evaluationmethodsmodeling_2013,costello_communityeffortassess_2014,kuffner_crowdsourcedanalysisclinical_2015,ewing_combiningtumorgenome_2015,hill_inferringcausalmolecular_2016}, FlowCAP \cite{aghaeepour_criticalassessmentautomated_2013,aghaeepour_benchmarkevaluationalgorithms_2016}, CASP \cite{moult_criticalassessmentmethods_2018,moult_criticalassessmentmethods_2016}, CAMI \cite{sczyrba_criticalassessmentmetagenome_2017}, Assemblathon \cite{earl_assemblathoncompetitiveassessment_2011,bradnam_assemblathonevaluatingnovo_2013}, MAQC/SEQC \cite{maqcconsortium_microarrayqualitycontrol_2006,shi_microarrayqualitycontrol_2010,seqc_comprehensiveassessmentrnaseq_2014}, and GA4GH \cite{krusche_bestpracticesbenchmarking_2019} consortia.

@@

By contrast, when introducing a new method, the focus of the benchmark will be on evaluating the relative merits of the new method. This may be sufficiently achieved with a less extensive benchmark, for example, by comparing against a smaller set of state-of-the-art and baseline methods. However, the benchmark must still be carefully designed to avoid disadvantaging any methods; for example, extensively tuning parameters for the new method while using default parameters for competing methods would result in a biased representation. Some advantages of a new method may fall outside the scope of a benchmark; for example, a new method may enable more flexible analyses than previous methods (e.g. beyond two-group comparisons in differential analyses \cite{weber_diffcytdifferentialdiscovery_2019}).

@@

The selection of methods to include in the benchmark will be guided by the purpose and scope of the study. A neutral benchmark should include all available methods for a certain type of analysis. In this case, the publication describing the benchmark will also function as a review of the literature; a summary table describing the methods is a key output (e.g. Figure 2 in \cite{saelens_comparisonsinglecelltrajectory_2019} or Table 1 in \cite{weber_comparisonclusteringmethods_2016}). Alternatively, it may make sense to include only a subset of methods, by defining inclusion criteria: for example, all methods that (i) provide freely available software implementations, (ii) are available for commonly used operating systems, and (iii) can successfully be installed without errors following a reasonable amount of trouble-shooting. Such criteria should be chosen without favoring any methods, and exclusion of any widely used methods should be justified. A useful strategy can be to involve method authors within the process, since they may provide additional details on optimal usage. In addition, community involvement can lead to new collaborations and inspire future method development. However, the overall neutrality and balance of the resulting research team should be maintained. Finally, if the benchmark is organized as a community challenge, the selection of methods will be determined by the participants. In this case, it is important to communicate the initiative widely -- for example, through an established network such as DREAM challenges. However, some authors may choose not to participate; a summary table documenting non-included methods should be provided in this case.

When developing a new method, it is generally sufficient to select a representative subset of existing methods to compare against. For example, this could consist of the current best-performing methods (if known), a simple baseline method, and any methods that are widely used. The selection of competing methods should ensure an accurate and unbiased assessment of the relative merits of the new approach, compared with the current state-of-the-art. In fast-moving fields, for a truly excellent benchmark, method developers should be prepared to update their benchmarks or design them to easily allow extensions as new methods emerge.

@@

Simulated data have the advantage that a known true signal (or ground truth) can easily be introduced; for example, whether a gene is differentially expressed. Quantitative performance metrics measuring the ability to recover the known truth can then be calculated. However, it is important to demonstrate that simulations accurately reflect relevant properties of real data, by inspecting empirical summaries of both simulated and real datasets (e.g. using automated tools \cite{soneson_unifiedqualityverification_2018}). The set of empirical summaries to use is context-specific; for example, for single-cell RNA-sequencing, dropout profiles and dispersion-mean relationships should be compared \cite{soneson_biasrobustnessscalability_2018}; for DNA methylation, correlation patterns among neighboring CpG sites should be investigated \cite{korthauer_detectionaccuratefalse_2019}; for comparing mapping algorithms, error profiles of the sequencing platforms should be considered \cite{caboche_comparisonmappingalgorithms_2014}. Simplified simulations can also be useful, to evaluate a new method under a basic scenario, or to systematically test aspects such as scalability and stability. However, overly simplistic simulations should be avoided, since these will not provide useful information on performance. A further advantage of simulated data is that it is possible to generate as much data as required; for example, to study variability and draw statistically valid conclusions.

Experimental data often do not contain a ground truth, making it difficult to calculate performance metrics. Instead, methods may be evaluated by comparing them against each other (e.g. overlap between sets of detected differential features \cite{nowicka_drimseqdirichletmultinomialframework_2016}), or against a current widely accepted method or gold standard (e.g. manual gating to define cell populations in high-dimensional cytometry \cite{weber_comparisonclusteringmethods_2016, aghaeepour_criticalassessmentautomated_2013}, or fluorescence in situ hybridization to validate absolute copy number predictions \cite{zheng_benchmarkingcontextsdetails_2017}). In the context of supervised learning, the response variable to be predicted is known in the manually labeled training and test data. However, individual datasets should not be overused, and using the same dataset for both method development and evaluation should be avoided, due to the risk of overfitting and overly optimistic results \cite{grimm_evaluationtoolsused_2015,jelizarow_overoptimismbioinformaticsillustration_2010}. In some cases, it is also possible to design experimental datasets containing a ground truth. Examples include: (i) spiking in synthetic RNA molecules at known relative concentrations \cite{jiang_syntheticspikeinstandards_2011} in RNA-sequencing experiments (e.g. \cite{seqc_comprehensiveassessmentrnaseq_2014,garalde_highlyparalleldirect_2018}), (ii) large-scale validation of gene expression measurements by quantitative polymerase chain reaction (e.g. \cite{seqc_comprehensiveassessmentrnaseq_2014}), (iii) using genes located on sex chromosomes as a proxy for silencing of DNA methylation status (e.g. \cite{law_voomprecisionweights_2014,fang_genomiclandscapehuman_2012}), (iv) using fluorescence-activated cell sorting to sort cells into known subpopulations prior to single-cell RNA-sequencing (e.g. \cite{soneson_biasrobustnessscalability_2018, schaum_singlecelltranscriptomics20_2018, zheng_massivelyparalleldigital_2017}), or (v) mixing different cell lines to create pseudo-cells \cite{tian_benchmarkingsinglecell_2019}. However, it may be difficult to ensure that the ground truth represents an appropriate level of variability -- for example, the variability of spiked-in material, or whether method performance on cell line data is relevant to outbred populations. Alternatively, experimental datasets may be evaluated qualitatively, for example, by judging whether each method can recover previous discoveries, although this strategy relies on the validity of previous results.

A further technique is to design semi-simulated datasets that combine real experimental data with an in silico (i.e., computational) spike-in signal; for example, by combining cells or genes from null (e.g. healthy) samples with a subset of cells or genes from samples expected to contain a true differential signal (examples include \cite{weber_diffcytdifferentialdiscovery_2019,arvaniti_sensitivedetectionrare_2017,rigaill_syntheticdatasets_2018}). This strategy can create datasets with more realistic levels of variability and correlation, together with a ground truth.

Overall, there is no perfect reference dataset, and the selection of appropriate datasets will involve tradeoffs, for example, regarding the level of complexity. Both simulated and experimental data should not be too simple (e.g. two of the datasets in the FlowCAP-II challenge \cite{aghaeepour_criticalassessmentautomated_2013} gave perfect performance for several algorithms) or too difficult (e.g. for the third dataset in FlowCAP-II, no algorithms performed well); in these situations, it can be impossible to distinguish performance. In some cases, individual datasets have also been found to be unrepresentative, leading to over-optimistic or otherwise biased assessment of methods (e.g. \cite{lowes_bralibasedentatale_2017}). Overall, the key to truly excellent benchmarking is diversity of evaluations, i.e., using a range of metrics and datasets that span the range of those that might be encountered in practice, so that performance estimates can be credibly extrapolated.

@@

There are three major strategies for choosing parameters. The first (and simplest) is to use default values for all parameters. Default parameters may be adequate for many methods, although this is difficult to judge in advance. While this strategy may be viewed as too simplistic for some neutral benchmarks, it reflects typical usage. We used default parameters in several neutral benchmarks where we were interested in performance for untrained users \cite{saelens_comparisonsinglecelltrajectory_2019,couronne_randomforestlogistic_2018,schneider_mortalityriskacute_2016}. In addition, for \cite{saelens_comparisonsinglecelltrajectory_2019}, due to the large number of methods and datasets, total runtime was already around a week using 192 processor cores, necessitating judgment in the scope of parameter tuning. The second strategy is to choose parameters based on previous experience or published values. This relies on familiarity with the methods and the literature, reflecting usage by expert users. The third strategy is to use a systematic or automated parameter tuning procedure -- for example, a grid search across ranges of values for multiple parameters or techniques such as cross-validation (e.g. \cite{saelens_comprehensiveevaluationmodule_2018}). The strategies may also be combined, for example, by setting non-critical parameters to default values and performing a grid search for key parameters. Regardless, neutrality should be maintained: comparing methods with the same strategy makes sense, while comparing one method with default parameters against another with extensive tuning makes for an unfair comparison.

For benchmarks performed to introduce a new method, comparing against a single set of optimal parameter values for competing methods is often sufficient; these values may be selected during initial exploratory work or by consulting documentation. However, as outlined above, bias may be introduced by tuning the parameters of the new method more extensively. The parameter selection strategy should be transparently discussed during the interpretation of the results, to avoid the risk of over-optimistic reporting due to expending more researcher degrees of freedom on the new method \cite{boulesteix_tensimplerules_2015,hu_parametertuningkey_2019}.

Software versions can also influence results, especially if updates include major changes to methodology (e.g. \cite{vaquero-garcia_leafcuttervsmajiq_2018}). Final results should generally be based on the latest available versions, which may require re-running some methods if updates become available during the course of a benchmark.

@@

Evaluation of methods will rely on one or more quantitative performance metrics (Figure  \ref{fig:fig2_metrics}A). The choice of metric depends on the type of method and data. For example, for classification tasks with a ground truth, metrics include the true positive rate (TPR; sensitivity or recall), false positive rate (FPR; 1 - specificity), and false discovery rate (FDR). For clustering tasks, common metrics include the F1 score, adjusted Rand index, normalized mutual information, precision, and recall; some of these can be calculated at the cluster level as well as averaged (and optionally weighted) across clusters (e.g. these metrics were used to evaluate clustering methods in our own work \cite{duo_systematicperformanceevaluation_2018,weber_comparisonclusteringmethods_2016}  and by others \cite{freytag_comparisonclusteringtools_2018,aghaeepour_criticalassessmentautomated_2013,wiwie_comparingperformancebiomedical_2015}). Several of these metrics can also be compared visually to capture the tradeoff between sensitivity and specificity, for example, using receiver operating characteristic (ROC) curves (TPR versus FPR), TPR versus FDR curves, or precision-recall (PR) curves (Figure \ref{fig:fig2_metrics}B). For imbalanced datasets, PR curves have been shown to be more informative than ROC curves \cite{saito_precisionrecallplotmore_2015,powers_visualizationtradeoffevaluation_2015}. These visual metrics can also be summarized as a single number, such as area under the ROC or PR curve; examples from our work include \cite{weber_diffcytdifferentialdiscovery_2019,soneson_biasrobustnessscalability_2018}. In addition to the tradeoff between sensitivity and specificity, a methods operating point' is important; in particular, whether the threshold used (e.g. 5\% FDR) is calibrated to achieve the specified error rate. We often overlay this onto TPR-FDR curves by filled or open circles (e.g. Figure \ref{fig:fig2_metrics}B, generated using the iCOBRA package \cite{soneson_icobraopenreproducible_2016}); examples from our work include \cite{weber_diffcytdifferentialdiscovery_2019,nowicka_drimseqdirichletmultinomialframework_2016,zhou_robustlydetectingdifferential_2014,soneson_differentialanalysesrnaseq_2016}.

@@

For methods with continuous-valued output (e.g. effect sizes or abundance estimates), metrics include the root mean square error, distance measures, Pearson correlation, sum of absolute log-ratios, log-modulus, and cross-entropy. As above, the choice of metric depends on the type of method and data (e.g. \cite{costello_communityeffortassess_2014,lindgreen_evaluationaccuracyspeed_2016}  used correlation, while \cite{moult_criticalassessmentmethods_2016} used root mean square deviation). Further classes of methods include those generating graphs, phylogenetic trees, overlapping clusters, or distributions; these require more complex metrics. In some cases, custom metrics may need to be developed (e.g. we defined new metrics for topologies of developmental trajectories in \cite{saelens_comparisonsinglecelltrajectory_2019}). When designing custom metrics, it is important to assess their reliability across a range of prediction values (e.g. \cite{gurevich_quastqualityassessment_2013,narzisi_comparingnovogenome_2011}). For some metrics, it may also be useful to assess uncertainty, for example, via confidence intervals. In the context of supervised learning, classification or prediction accuracy can be evaluated by cross-validation, bootstrapping, or on a separate test dataset (e.g. \cite{gardner_identifyingaccuratemetagenome_2019,aghaeepour_benchmarkevaluationalgorithms_2016}). In this case, procedures to split data into training and test sets should be appropriate for the data structure and the prediction task at hand (e.g. leaving out whole samples or chromosomes \cite{schreiber_pitfallmachinelearning_2019}).

Additional metrics that do not rely on a ground truth include measures of stability, stochasticity, and robustness. These measures may be quantified by running methods multiple times using different inputs or subsampled data (e.g. we observed substantial variability in performance for some methods in \cite{soneson_biasrobustnessscalability_2018,weber_comparisonclusteringmethods_2016}). Missing values may occur if a method does not return any values for a certain metric, for example, due to a failure to converge or other computational issues such as excessive runtime or memory requirements (e.g. \cite{saelens_comparisonsinglecelltrajectory_2019, soneson_biasrobustnessscalability_2018,weber_comparisonclusteringmethods_2016}). Fallback solutions such as imputation may be considered in this case \cite{bischl_benchmarkinglocalclassification_2013}, although these should be transparently reported. For non-deterministic methods (e.g. with random starts or stochastic optimization), variability in performance when using different random seeds or subsampled data should be characterized. Null comparisons can be constructed by randomizing group labels such that datasets do not contain any true signal, which can provide information on error rates (e.g. \cite{weber_diffcytdifferentialdiscovery_2019,zhou_robustlydetectingdifferential_2014,law_voomprecisionweights_2014}). However, these must be designed carefully to avoid confounding by batch or population structure, and to avoid strong within-group batch effects that are not accounted for.

For most benchmarks, multiple metrics will be relevant. Focusing on a single metric can give an incomplete view: methods may not be directly comparable if they are designed for different tasks, and different users may be interested in different aspects of performance. Therefore, a crucial design decision is whether to focus on an overall ranking, for example, by combining or weighting multiple metrics. In general, it is unlikely that a single method will perform best across all metrics, and performance differences between top-ranked methods for individual metrics can be small. Therefore, a good strategy is to use rankings from multiple metrics to identify a set of consistently high-performing methods, and then highlight the different strengths of these methods. For example, in \cite{weber_comparisonclusteringmethods_2016}, we identified methods that gave good clustering performance, and then highlighted differences in runtimes among these. In several studies, we have presented results in the form of a graphical summary of performance according to multiple criteria (examples include Figure 3 in \cite{saelens_comparisonsinglecelltrajectory_2019} and Figure 5 in \cite{soneson_biasrobustnessscalability_2018} from our work; and Figure 2 in \cite{sage_quantitativeevaluationsoftware_2015} and Figure 6 in \cite{korthauer_practicalguidemethods_2019} from other authors). Identifying methods that consistently underperform can also be useful, to allow readers to avoid these.
